### 2012-07-18 Doc/VRIPHYS2012/vriphys2012.tex
modifs in intro
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@428 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\title[Short title]%
      {Fill in the Title} % TODO

@@

Objects with thin structure occur in many forms e.g. textiles, paper,
leaves etc. Such objects are also very common in anatomical structures for
example in the form of tubular objects like blood vessels or colon, in the
form of thin membranes, but also many others (lens, skin, \ldots). Because
they have high ratio between width and thickness methods for volumetric
deformation modeling cannot be immediately used. For this reason they form
a very specific field of research.

Methods based on continuum mechanics and namely theory of elasticity gained
on popularity in computer graphics and real-time simulators ever since
Terzopoulos \cite{Terzopoulos1987} presented his work on elastic
deformation modeling. While the methods using finite elements and theory of
elasticity are fairly popular in volumetric modeling \footnote{TODO: do we
need ref. here?} of deformations, they are still frowned upon in the area
of thin structures. Instead either the mass-spring models \cite{Volino2009}
or bending models \cite{Grinspun2003,Choi2007} are preferred. Unfortunately
they have the disadvantage that physical parameters of a material (like
Young's modulus and Poisson's ratio) cannot be used directly.

The usual argument against FEM based methods is that they are
computationally too expensive. This is indeed true if lots of elements are
used to discretize the object or non-linear models are used. However,
exploiting the bending property of the shell elements allows us to achieve
good results even with relatively small number of elements. We can use
high-polygonal mesh only for rendering or collision detection and response.
Similar idea has been already presented in the work of Bouthors et al.
\cite{Bouthors2007}. They used twinned meshes for animation of mesh defined
by implicit surface. Simple low-polygonal mesh was used for mechanical FEM
model and the mesh was subdivided in areas with high curvature after every
step of the simulation to produce finer mesh for rendering. The underlaying
FEM model is however not aware of the implicit surface used. We build upon
the work of Comas \cite{Comas2010c} who used a simple cubic polynomial to
describe the element geometry and derived the shell model from it. Comas's
element has several drawbacks that we needed to address. Namely it produces
discontinuities between the elements on the edges, it is missing one
rotational degree of freedom and is not completely symmetric in context of
deformations.

% TODO: integrate better to the rest of the section
Ubach and Oñate \cite{Ubach2010} also present a shell element based on
Bézier triangles. The element is however free from rotational degrees of
freedom which is unsuitable for physical simulators. It also reduces the
sparsity of stiffness matrix because of increased interconnection between
nodes.

% In this paper we present a new shell element based on the
% formulation of Bézier triangle.
%
% TODO: paper structure
TODO: paper structure

>>>>

\title[Short title]%
      {Bezier shell finite element for interactive medical simulation} % TODO

@@

Thin objects are very frequent in our every-day life. % thin wall objects ?
%Objects with thin structure occur in many forms 
e.g. textiles, paper, leaves etc. 
Such objects are also very common in anatomical or pathological structures for example tubular structures (like blood vessels, colon \ldots), thin bag structures (%vessie, 
Glisson capsule, aneurysms,  \ldots), and also many others (eyes, skin, \ldots). 
In the field of surgical simulation, we try to reproduce or anticipate the mechanical behavior of these structures during the surgery.
Thus, it is very important to capture accurately the physical behavior of these structure.
However, in the same time, it is also very important to maintain fast computation rates, in order to be compatible with the targeted applications (interactive simulation for education and for planning).

This problem has been faced by the computer graphics community but for different applications.
For instance, many related work faces the problem of the simulation of cloth using mass-spring models \cite{Volino2009} or bending models \cite{Grinspun2003,Choi2007}. %% Todo verify.... 
However, these models are discrete so their behavior depends on the mesh and their parameters are not easily measurable.
This is completely acceptable for some computer graphics application, where the parameters can be tuned to obtain a realistic animation.
In the case of medical simulation, more physics-based modeling is needed.

Methods based on continuum mechanics and namely theory of elasticity gained on popularity in computer graphics and real-time simulators ever sinceTerzopoulos \cite{Terzopoulos1987} presented his work on elastic deformation modeling.  %% Do we
While the methods using finite elements and theory of elasticity are fairly popular in volumetric modeling \footnote{TODO: do we need ref. here?  yes but for medical applications} of deformations, they are still frowned upon in the area of thin structures. 
Yet, a special field of continuum mechanics has studied the deformations of these objects, and has proposed models based on shell theory.
This theory can be integrated using Finite Elements, but there is two drawbacks: (i) the continuum equations requires square spatial derivations that can not be handled with linear interpolation and (ii) the final system of equations that describes the mechanics is non-linear (if the deformations are not small).

The usual argument against non-linear FEM based methods is that they are computationally too expensive. 
This is indeed true if lots of elements are used to discretize the object.
A classical strategy, is to use high-polygonal mesh for rendering or collision detection and response, while maintaing a reasonably low number of elements for the mechanics.
The work of Bouthors et al.\cite{Bouthors2007} is a good application example of this strategy for surface deformations.
But this strategy is not sufficient in our case because the problem of performance is also greatly impacted when using elements that have an interpolation with high order polynoms.
This type of interpolation is necessary because bending continuum model (based on Kirchhoff-Love theory for thin plates)  requires the computation of square spatial derivatives.
In \cite{Comas2010c}, Comas and al.  use an hybrid interpolation: linear interpolation is used for in-plane deformations and cubic polynomials for bending deformations.
The element is piloted by 3 nodes on which 6 degrees of freedom are defined. 
From this point of view, our work is derived from this approach but we address several noted drawbacks: the formulation misses one rotational degree of freedom, the cubic interpolation is not symmetric and produces $C^0$  discontinuities between elements on edges.
Ubach and al. \cite{Ubach2010}  present a shell element based on B\'ezier triangles that allows for a better behavior.
However, the B\'ezier interpolation defined on each triangle element is based on the motion computed on adjacent triangles.
So the stiffness forces computed on each element not only depends on nodes of the element but also on all the adjacent nodes. 
This greatly reduces the sparsity of the global stiffness matrix and consequently increases the computation time.

hell element based on the
% formulation of Bézier triangle.
%
% TODO: paper structure
TODO: paper structure
